\chapter{Introduction}
\label{chapter:introduction}
Software engineering is becoming an increasingly challenging task nowadays. Developing software with complex architectures and nontrivial implementations is a prevalent task for the modern software engineer, having implications on how software is developed. At the same time, software that can be proven to be error-free has become increasingly important. The reason for this is apparent. Sophisticated systems automate more and more crucial tasks. Failure of these systems might have enormous consequences, especially for safety-critical and healthcare systems. Therefore, multiple strategies have been developed over the years to ensure that crucial parts of these systems are error-free.

An increasingly popular method for dealing with the development of complex systems is by using domain-specific models. Model-Driven Engineering (MDE) is a field within software engineering that focuses on using and creating domain models that describe complex software systems on a domain level. These models can then be used for different tasks, depending on the type of model. These tasks include code generation, but also different forms of verification of the software. By using these models, it becomes easier to reason about the developed software, while also allowing for systematic code generation and verification.

Although domain-specific models provide a strategy to deal with the development of complex systems, they do not automatically ensure that the software is error-free. Software verification is an essential strategy in ensuring that software systems are error-free. Modern methods of software verification use automated tools that can use software models to verify the correctness of a system. These tools use some model of the system to verify a set of requirements provided by the software engineer. By using structural checks on the model, the tool can tell if these requirements are met.

A possible problem that might arise when using domain-specific models for software development is the interoperability of different models. Within the area of MDE, a lot of different frameworks and tools exist. Each of these frameworks and tools focuses on a specific set of functionality. As a result, models created in one framework well suited for code generation, might not be useful in the context of software verification. In an ideal world, the format of the produced models would be standardised across all frameworks for smooth interoperability. In reality, different frameworks use different formats which are optimised for their specific set of functions. These different formats make it difficult to share models across different frameworks and applications.

Model transformation is a concept in the field of MDE that focuses on solving this problem. Model transformation is an automated way of modifying and creating models by transforming existing models. By using model transformations, it is possible to transform a model that is tailored towards code generation into a model that is suited for software verification, without the need to create a new model for this purpose.

Model transformations have already led to various tools and services that can export and import models in different tools and frameworks. These tools and services allow a software engineer to transform a model suited for code generation into a model suited for verification, and therefore use one model and its transformations to achieve both tasks. Sadly, these model transformations rarely have a formal foundation. Having a formal foundation for the model transformations is useful in the context of software verification since it allows for proving the correctness of the transformation itself. When a transformed model is used to verify a software system, the results of the verification can only be considered correct if the transformation is correct. Without proof of correctness of the transformation, it might be that the verification results are incorrect because the original model might have a different meaning than the transformed model.

This thesis will contribute to fields of MDE and Software Verification by specifying a formal foundation for model transformations between EMF/Ecore (\cref{sec:background:eclipse_modeling_framework}), a framework for software modelling in which various models can be created, and GROOVE (\cref{sec:background:groove}), a tool for software verification based on graph grammars. Furthermore, a framework is presented in which these model transformations can be proven correct, allowing the user to build correct model transformations iteratively.

\input{tex/01_introduction/01_formalisation_of_model_transformations.tex}
\input{tex/01_introduction/02_correctness_of_model_transformations.tex}
\input{tex/01_introduction/03_approach.tex}
\input{tex/01_introduction/04_research_question.tex}
\input{tex/01_introduction/05_validation.tex}
\input{tex/01_introduction/06_related_work.tex}
\input{tex/01_introduction/07_contribution.tex}
\input{tex/01_introduction/08_outline.tex}