\chapter{Conclusion}
\label{chapter:conclusion}

In \cref{chapter:introduction} of this thesis, the need for verification in complex software projects was discussed. Modern approaches to software verification use models to verify a piece of software. However, not all models are suited for this task. Moreover, creating multiple models in different modelling languages for the same software is time-consuming and expensive. The use of model transformations to automatically transform models between different modelling languages was presented as a solution. However, in order for model transformations to be of use in the context of software verification, a formal foundation for the model transformations is required. This thesis has shown a formal foundation for model transformations between EMF/Ecore and GROOVE, which can be of use in the field of software verification.

As discussed in \cref{sec:background:eclipse_modeling_framework}, EMF/Ecore is a popular modelling language for modelling software. However, its models are not well-suited for verification. As discussed in \cref{sec:background:groove}, GROOVE is created especially for software verification but uses a different modelling language than EMF/Ecore. Therefore, model transformations should be used to convert between the two models automatically. A formalisation of the modelling languages themselves was provided in \cref{chapter:formalisations}. These formalisations are the foundation of the formalisation of the model transformations presented as part of the transformation framework.

The transformation framework presented in \cref{chapter:transformation_framework} provides the main result of this thesis. The transformation framework includes a formalisation of transformation functions between Ecore models and GROOVE graphs, which allows the user to reason about these transformation functions formally. Furthermore, the transformation framework has presented a structured way to compose these transformation functions iteratively while maintaining the correctness. The composability of these transformation functions is essential, as it allows the user of the framework to build significant model transformations without loss of correctness. The correctness property is relevant in the field of software verification, as it allows for verification of Ecore models within GROOVE, without the loss of confidence that the results might be incorrect due to transformation errors.

In order to further validate the transformation framework, \cref{chapter:library_of_transformations} has presented a small library of transformations that can be used within the transformation framework. The transformations from the library presented in this chapter are all small transformations with a few elements, that can be added to a larger model using the framework. In this way, the transformations can be used to compose significant transformations. The library of transformations allows a user to build specific models without the need to define transformation functions for each created model.

Finally, \cref{chapter:application} has shown an application of the transformation framework and the library of transformations in a practical example. Throughout the chapter, a single model is built from scratch, showing each of the steps taken to build the model using the transformation framework. Each step combines a transformation from the library of transformations with the existing transformation function, making the transformation function larger and allowing more complex models to be transformed.

In this final chapter, the work presented by the thesis will be concluded. First, the advantages and limitations of the work will be discussed. Then the work will be evaluated and the research questions introduced in \cref{sec:introduction:research_question} will be answered. Finally, some proposals for future work are discussed.

\input{tex/07_conclusion/01_limitations.tex}
\input{tex/07_conclusion/02_evaluation.tex}
\input{tex/07_conclusion/03_future_work.tex}